---
title: "Stat 541 Final Portfolio"
format: pdf
editor: visual
self-contained: true
toc: true
---

```{r}
#| include: false
#| label: packages-portfolio-objectives

library(tidyverse)

portfolio_levels <- tibble(
  Objective = character(), 
  Level = numeric() 
  )
```

# Independent Learning (IL):

*These objectives show your ability to seek out new information and adapt to new tools to solve data analysis problems.*

------------------------------------------------------------------------

## \[IL-1\] Adding new skills:

-   I can find and adopt new packages to accomplish tasks.
-   I can adapt to different syntax styles (tidy, base, formula style, data.table).
-   I can use tutorials, etc. to enhance my understanding of new concepts

```{r}
#| include: false
#| label: IL-1

## Fill in the line below with your self-assigned level for this objective.
## Remember, you have the following levels to choose from:

# 1 = novice
# 2 = proficiency
# 3 = mastery 
# 4 = expert

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "IL-1",
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[IL-2\] Online resources:

-   I can use online resources (Google, ChatGPT, StackOverflow) to solve problems, debug, or find new tools.
-   I can find source code for similar projects to use as starting points for my own
-   I can read the documentation of an API to figure out how to access data.

```{r}
#| include: false
#| label: IL-2

## Fill in the line below with your self-assigned level for this objective.
my_level <- 3

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "IL-2", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

For the Week 3 Shiny Project, I found the base code for our server, ui, and global files on [RStudios gallery](https://shiny.posit.co/r/gallery/start-simple/word-cloud/) of Shiny apps. Furthermore, I relied on ChatGTP to support some of my changes to modifying the base code. For example, I noticed that my word cloud plot cut off words when the size or amount of words exceeded some limit. I queried "my plot gets cutt off when too many words are allowed in my word cloud" after providing my code, and ChatGPT provided three solutions. I recognized that one of them was impossible, and after testing the remaining two found that only one of them solved my issue, but solve it did!

------------------------------------------------------------------------

# Reproducible Workflow (RW):

*These objectives show your ability to produce artifacts and deliverables that are organized, documented, version tracked, and responsibly designed.* \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## \[RW-1\] File, code, and data management:

-   I can use Git and GitHub to track my progress and collaborate (creating repos, cloning, forking, pull requesting).
-   I always use R Projects and the {here} package to organize my scripts, notebooks, data, and applications.

```{r}
#| include: false
#| label: RW-1

my_level <- 3

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "RW-1",
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

Look no further than this portfolios [github page](https://github.com/Wander03/Stat_551_portfolio) itself! Here you can see evidence of creating my own repo, then [forking and pull requesting](https://github.com/Wander03/Stat_551_portfolio/network) in a neat, orderly, manner. A beautiful example of cloning occurred during the [Shiny Project](https://github.com/keasley8/Project_1_Shiny), which is wonderfully organized into a "data" folder and a "app" folder housing our R scripts (named fake_news). Within this folder itself you can find examples of the {here} package being used within the [global notebook](https://github.com/keasley8/Project_1_Shiny/blob/main/fake_news/global.R), aiming at the "data" folder and then the necessary data set file.

------------------------------------------------------------------------

## \[RW-2\] Notebooks:

-   I can use Quarto and/or R Markdown to produce a reproducible notebook and polished rendered document.
-   I can use appropriate chunk options (echo, error, cache, etc.) to render my qmd/Rmd quickly and cleanly.

```{r, include = FALSE}
#| include: false
#| label: RW-2

my_level <- 3

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "RW-2",
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

The [practice lab](https://github.com/Cal-Poly-Advanced-R/practice-lab-Wander03/blob/master/practice-lab.qmd) is the all-in-one assignment with a variety of different chunk options decorating the Quarto file. From echo, to error, to message, it has it all! Although it has wonderful labels as well, lets look to lab 1 for examples of this. Specifically, [part 2](https://github.com/Cal-Poly-Advanced-R/lab-1-review-Wander03/blob/master/part2.qmd) does a perfect job with the label chunk option naming each chunk appropriately (it even has warning and message options added as needed). Additionally, we can see the [rendered output for part 2](https://github.com/Cal-Poly-Advanced-R/lab-1-review-Wander03/blob/master/part2.html), and part 1 if you want it, presenting the fruits of our labor (minus the chunk labels, those are hidden away).

------------------------------------------------------------------------

## \[RW-3\] Code style

-   My code is clear, readable, well-organized, and well-commented.
-   I can use a package-based workflow to organize my analyses

```{r}
#| include: false
#| label: RW-3

my_level <- 4

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "RW-3",
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

My code, especially my functions, include comments (almost) line by line, as well as an overall comment describing the inputs and outputs of the function as a whole. A fresh example (at the time of writing this) is from lab 4 (where did lab 3 go? who knows!). Within this [pull request](https://github.com/Cal-Poly-Advanced-R/lab-3-using-apis-Ryan-DeStefano/pull/6/files), lines 65 to 155 contain parts of a helper function I wrote. The aforementioned pull request shows my addition of comments so my group mates, and others, can understand the logic of my function. Later on in a [separate pull request](https://github.com/Cal-Poly-Advanced-R/lab-3-using-apis-Ryan-DeStefano/pull/7/commits/f627bd450f2da77f9de80d84c823b40614d3894e) I added an overall header to said function!

------------------------------------------------------------------------

# Technical Communication (TC):

*These objectives show your ability to communicate the processes you have implemented in your code, as well as the data conclusions and results.*

------------------------------------------------------------------------

## \[TC-1\] Project summaries:

-   I can clearly and succinctly summarize the contributions of my project.
-   I accurately interpret statistical or modeling results.
-   I consider the appropriate scope and impact of my project results.

```{r}
#| include: false
#| label: TC-1

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "TC-1", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[TC-2\] Documentation:

-   I provide ample documentation and tutorials for my custom functions.
-   I provide user-friendly guides for my tools and software

```{r}
#| include: false
#| label: TC-2

my_level <- 4

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "TC-2", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

For Lab 1, I was the sole contributor to Part 1, the [Q-Q Plot custom function](https://github.com/Cal-Poly-Advanced-R/lab-1-review-Wander03/pull/3#issue-2228813761). This involved creating a custom Q-Q Plotting function and testing it with the cars dataset. My function not only includes comments detailing what each of the following line(s) of code do, but additionally has an [overview comment](https://github.com/Cal-Poly-Advanced-R/lab-1-review-Wander03/pull/5). The overview comment lists what data types each input is required to be along with what the output of the function is. For another example, see my [get_passes function from lab 4](https://github.com/Cal-Poly-Advanced-R/lab-3-using-apis-Ryan-DeStefano/blob/master/main.qmd) (as mentioned in RW-3).

------------------------------------------------------------------------

# Data Manipulation (DM):

*These objectives relate to the collection, cleaning, processing, and preparing of datasets for analysis.*

------------------------------------------------------------------------

## \[DM-1\] Data preparation:

-   I can read in datasets to R, including untidy ones.
-   I can clean datasets to deal with missing data, typos, poor formatting, etc.

```{r, include = FALSE}
#| include: false
#| label: DM-1

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "DM-1", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[DM-2\] Data wrangling

-   I can cleverly use pivoting, grouping, and joining to wrangle data.
-   I can use mapping ({purrr}), applying (tapply, lapply, ...), and/or iteration (for loops) to perform repeated tasks.

```{r, include = FALSE}
#| include: false
#| label: DM-2

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "DM-2",
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[DM-3\] Data formats

-   I can use API urls to access JSON data and convert it to a data frame

```{r, include = FALSE}
## Fill in the line below with your self-assigned level for this objective.
my_level <- 0

portfolio_levels <- portfolio_levels %>%
  bind_rows(tibble(Objective = "DM-3", 
                   Level = my_level))
```

### Level: `r my_level`

### Justification

------------------------------------------------------------------------

## \[DM-4\] Data collection

-   I can webscrape simple tables and information

```{r}
#| include: false
#| label: DM-4

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "DM-3", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification

------------------------------------------------------------------------

# Professional Visualization (PV):

------------------------------------------------------------------------

## \[PV-1\] ggplot: grammar of graphics

-   I can use less common geometries, including those from ggplot extension packages.
-   I can use the correct aesthetics to map variables
-   I understand how geometries inherit aesthetics I can add annotations to my plot

```{r}
#| include: false
#| label: PV-1

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "PV-1", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[PV-2\] ggplot: theme

-   I can edit the titles, subtitles, captions, axis labels, etc. to create a clearly labelled plot
-   I can choose colors ("scales") and themes to make a visually pleasing and accessible plot

```{r}
#| include: false
#| label: PV-2

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "PV-2", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[PV-3\] Dynamic visualizations

-   I can use a package like {gganimate} to create self-contained gifs
-   I can use a package like {plotly}, {ggplotly}, {leaflet}, {ggirafe}, etc. to make interactive html widgets

```{r}
#| include: false
#| label: PV-3

my_level <- 3

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "PV-3", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

\[ADD JUSTIFICATION FOR GGANIMATE WHEN YOU USE IT (I KNOW YOU WILL USE IT FUTURE ME\]

For Lab 2, although I created the 2nd Visualization, I was the first to [create my graph](https://github.com/Cal-Poly-Advanced-R/lab-2-visualization-kchan78/pull/12/commits/d43d83c056aa80774ebe7184c7967f9eadcd7f37). For my team I pioneered the use of {ggirafe}, picking out how to create pop-up labels when hovering over data points and format them nicely. Looking at Katherines contributes later, you can see the same general formatting for [her use](https://github.com/Cal-Poly-Advanced-R/lab-2-visualization-kchan78/pull/18/commits/cd8b30c4b1b6a1803994a51dd4bd50f07a5d5c6d) of {ggirafe}. Additionally, I [improved upon the leaflet](https://github.com/Cal-Poly-Advanced-R/lab-2-visualization-kchan78/pull/23/commits/92e2348539c67414c64048518dffebfdfd1fb5bb) Kai created, adding labels on hover and setting the initial View.

------------------------------------------------------------------------

## \[PV-4\] Shiny

-   I can create a functional Shiny app.
-   I understand the principle of reactivity, and how to use it.

```{r}
#| include: false
#| label: PV-4

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "PV-4", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

# Software Development (SD):

*These objectives relate to your ability to develop correct, usable, well-designed, and sophisticated software in the R language.*

------------------------------------------------------------------------

## \[SD-1\] R programming language details

-   I understand non-standard evaluation (aka "Tidy Eval" or "unquoted objects"), and I can use tunneling in my functions.
-   I understand functional programming, and I can use functions as objects in my code design
-   I understand object-oriented programming, and I can define my own classes and methods.

```{r}
#| include: false
#| label: SD-1

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "SD-1", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[SD-2\] Package creation:

-   I can create a folder that is installable as an R package, possibly using {usethis} helper functions
-   I can document my functions using {roxygen2} style commenting
-   I can write and run unit tests using {testthat}
-   I can design a package that is user-friendly and has well-designed functions.

```{r}
#| include: false
#| label: SD-2

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "SD-2", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification

------------------------------------------------------------------------

# Matrix Operations (MO):

*These objectives show your ability to manipulate data-related information in the form of vectors and matrices, rather than in high-level data structures.*

------------------------------------------------------------------------

## \[MO-1\] Theory:

-   I understand the difference between ordinary multiplication and matrix multiplication, and how to implement each in R
-   I can implement and briefly explain the matrix equations for multiple linear regression

```{r}
#| include: false
#| label: MO-1

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "MO-1", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[MO-2\] Object structures:

-   I can convert data objects into the necessary matrix structures to perform operations on them.
-   I can convert results of matrix operations to convenient data analysis formats

```{r}
#| include: false
#| label: MO-2

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "MO-2", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

# Algorithms and Iteration (AI):

*These objectives ask you to design code-based approaches to statistical computing problems, usually involving iteration to a stopping condition.*

------------------------------------------------------------------------

## \[AI-1\] Creating an algorithm:

-   I can invent and implement my own iterative algorithm.

```{r}
#| include: false
#| label: AI-1

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "AI-1", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[AI-2\] Generative art:

-   I can apply a variety of generative art functions to make a visually pleasing piece.
-   I can explain why particular changes to the code result in particular differences in the visualization.

```{r}
#| include: false
#| label: AI-2

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "AI-2", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

# Code Design (CD):

*These objectives relate to making wise or clever choices in how you implement a procedure in code; including creating functions and objects, or thinking about the clarity and efficiency of processes.*

------------------------------------------------------------------------

## \[CD-1\] Speed and Efficiency:

-   I can recognize moments of possible slowdown in my code, and use built-in functions or parallelizing to speed them up.
-   I always use and design vectorized functions whenever possible.

```{r}
#| include: false
#| label: CD-1

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "CD-1", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[CD-2\] Object handling:

-   I can make reasonable choices in my code design about when to save intermediate objects.
-   I can convert objects between types and structures as needed.

```{r}
#| include: false
#| label: CD-2

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "CD-2", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[CD-3\] Supporting functions:

-   I write helper/shortcut functions to streamline repeated tasks and make my code easier to read.
-   I use intermediate functions to streamline repeated or looping processes.

```{r}
#| include: false
#| label: CD-3

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "CD-3", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

## \[CD-4\] Algorithmic process:

-   My loops are clean and efficient
-   Proper values are calculated to update objects and/or determine stopping conditions
-   I have built in checks for possible problems or extreme cases in the algorithm

```{r}
#| include: false
#| label: CD-4

my_level <- 1

portfolio_levels <- portfolio_levels %>%
  bind_rows(
    tibble(Objective = "CD-4", 
           Level = my_level)
    )
```

### Level: `r my_level`

### Justification:

------------------------------------------------------------------------

# Summary

```{r}
#| echo: false
#| label: summary-graph

portfolio_levels %>%
  mutate(
    Category = str_extract(Objective, "[A-Z]{2}"),
    Level = factor(Level)
  ) %>%
  ggplot(
    aes(x = Level, 
        y = (..count..)/sum(..count..)
        )
    ) +
  geom_bar() +
  labs(x = "Level",
       y = "Percent") +
  theme_bw() +
  scale_y_continuous(labels = scales::percent)
```

## Grade

Based on the summary plot above, I believe I have earned a \_\_\_ in STAT 541.

### Justification
